Automated iterative near-term forecasting for the Portal Project

### 2018-02-20 manuscript/portal_iterative_forecasting_ms.Rmd
Fix broken references

---
title: "Automated iterative near-term forecasting for the Portal Project"
author:
- Ethan P. White$^{1,2,3}$
- Glenda M. Yenni$^1$
- Shawn D. Taylor$^4$
- Erica M. Christensen$^1$
- Ellen K. Bledsoe$^4$
- Juniper L. Simonis$^1$
- S. K. Morgan Ernest$^{1,3}$
csl: mee.csl
output:
  pdf_document:
    keep_tex: yes
  html_document: default
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
header-includes:
- \usepackage{times}
- \usepackage{setspace}
- \usepackage{booktabs}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
fontsize: 12pt
bibliography: refs.bib
---
\raggedright

@@

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of how well models capture the processes governing ecological systems [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology. Since then, an increasing number of ecological forecasts are being published. Most of these forecasts, however, are made once, published, and never assessed or updated. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build. Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petch2015; @dietze2018]. More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

@@

Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise. It is rare to find such breadth within a single research group, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab. When teams have a breadth of expertise, communication can be challenging [@winowiecki2011]. We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group. The two labs are part of a joint interdisciplinary ecology group that has a mission of breaking down barriers between field and computational/theoretical ecologists  (http://weecology.org). Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and informal lab training efforts. In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions. This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions. Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs. This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves. To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future  [@read2017].

>>>>

---
title: "Automated iterative near-term forecasting for the Portal Project"
author:
- Ethan P. White$^{1,2,3}$
- Glenda M. Yenni$^1$
- Shawn D. Taylor$^4$
- Erica M. Christensen$^1$
- Ellen K. Bledsoe$^4$
- Juniper L. Simonis$^1$
- S. K. Morgan Ernest$^{1,3}$
csl: mee.csl
output:
  html_document: default
  pdf_document:
    keep_tex: yes
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
header-includes:
- \usepackage{times}
- \usepackage{setspace}
- \usepackage{booktabs}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
fontsize: 12pt
bibliography: refs.bib
---
\raggedright

@@

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of how well models capture the processes governing ecological systems [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology. Since then, an increasing number of ecological forecasts are being published. Most of these forecasts, however, are made once, published, and never assessed or updated. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build. Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petchey2015; @dietze2018]. More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

@@

Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise. It is rare to find such breadth within a single research group, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab. When teams have a breadth of expertise, communication can be challenging [@winowiecki2011]. We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group. The two labs are part of a joint interdisciplinary ecology group that has a mission of breaking down barriers between field and computational/theoretical ecologists  (http://weecology.org). Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and informal lab training efforts. In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions. This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions. Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs. This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves. To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future  [@read2016].
